---
title: "p8105_hw2_yc3242"
author: "Youn Kyeong Chang (uni# yc3242)"
date: "October 3, 2018"
output: github_document
---

I used the tidyverse library through the whole problem set.

```{r}
library(tidyverse)
```


# Problem 1

First, I loaded and cleaned dataset for problem 1. 

```{r data import}
nyc_data = 
  read_csv(file = "./hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

This NYC Transit data (`nyc_data`) is composed of *line, station name, its latitude and longitude, routes it serves, entrance type, whether or not you can enter the station, presence of ticket vending machine*, and *ADA compliance*. These variables were selected from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv` original dataset which was followed by `clean_names()` function in the janitor library. After selecting variables process, `entry` variable was converted from character variable to logical variable. The size of the NYC Transit data is `r nrow(nyc_data)` rows and `r ncol(nyc_data)` columns. These data are not tidy because columns are values not variables and there are many duplicated rows. 

To remove duplicated rows, I used `distinct()` function as below:

```{r}
nyc_data = distinct(nyc_data)
```

The **size** of the NYC Transit data is now `r nrow(nyc_data)` rows and `r ncol(nyc_data)` columns.

### Questions

**1.1. How many distinct stations are there?**

```{r distinct_stations}
nyc_data %>% 
  distinct(line, station_name, .keep_all = TRUE) %>% 
  count()
```

There are **465** distinct stations.


**1.2. How many stations are ADA compliant?**

```{r ada}
nyc_data %>% 
  distinct(line, station_name, .keep_all = TRUE) %>% 
  filter(ada == TRUE) %>% 
  count(ada)
```

There are **84** ADA compliant stations.


**1.3. What proportion of station entrances / exits without vending allow entrance?**

```{r proportion}
nyc_data %>% 
  filter(vending == "NO") %>% 
  summarise(mean(entry == TRUE))
```

The proportion of station entrances / exits without vending allow entrance is **0.385**.


**1.4. How many distinct stations serve the A train?**

To answer the questions 1.4. and 1.5., I reformatted data as below so that route number and route name are distinct variables.

```{r reformat}
reform_nyc_data = 
  nyc_data %>%
  gather(key = "route", value = "route_name", route1:route11) %>% 
  separate(route, into = c("remove", "route_number"), sep = 5) %>% 
  select(-remove)
```

And then, I calculated distinct stations serve the A train.
```{r distinct_A}
reform_nyc_data %>% 
  distinct(line, station_name, route_name) %>%
  filter(route_name == "A") %>%
  count()
```

There are **60** distinct stations serve A train.


**1.5. Of the stations that serve the A train, how many are ADA compliant?**

```{r ada_A}
reform_nyc_data %>% 
  distinct(line, station_name, route_name, .keep_all = TRUE) %>%
  filter(route_name == "A", ada == "TRUE") %>%
  count()
```

There are **17** ADA compliant distinct stations serve A train.



